Refunctionalization of Abstracting Abstract Machines (Functional Pearl)

### 2018-03-16 paper.tex
Reworking contributions+outline

\subsection{Contributions}

@@

We summarize our contributions as follows:

\begin{itemize}
\item The main contribution of this paper is the filling of the gap
  between small-step AAM and big-step ADI by developing a series of
  systematic transformations.
  We show that the correspondence between abstract
  machines and interpreters exists not only in concrete semantics artifact,
  but also in abstract semantics artifact.
  Galois connection?
  \note{JD: Section number?}

\item A new understanding of pushdown control-flow analysis by
  refunctionalization.
  We present an abstract interpreter obtained from refunctionalizing
  a small-step abstracting abstract machine.
  This abstract interpreter naturally exhibits the pushdown control-flow property,
  as the refunctionalization forces us to sequentialize the
  abstract execution.
  We also analyze its correctness, soundness, and termination.

\item Further, we show that defunctionalization and refunctionalization
  of abstract interpreters play important roles for the call stack of the analyzed language.
  Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
  control-flow property (without extra efforts), as well as explain
  why ADI naturally extends pushdown control-flow property from its
  defining language.

\subsection{Outline}

\note{TR: can this be merged with contributions? (they currently lack section refs)}
The remainder of this paper is organized as follows.
We begin by describing the target language we will analyze in Section~\ref{background},
including a concrete CESK machine as small-step operational semantics for it.
We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
nondeterministic choices of closures to another layer of continuations,
therefore linearizing the abstract execution.
Section~\ref{fusing} then discusses fusing the driver and step functions to a single
function.
We further identify in Section~\ref{disen} the first-order data type representations of
continuations, and disentangle the code blocks for dispatching data types to separate
functions.
Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
In this section, the continuations represented by data types become higher-order
functions.
In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
different direct style, one is implemented with monads, the other is using
delimited controls.

Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.

>>>>

\note{rest of intro goes here}

The rest of this paper is structured as follows: we begin by reviewing some background
necessary for this work in Section~\ref{background}, as well as introducing some of the basic
code structures used throughout the paper. We then address the main contribution of this
paper, which is the filling of the gap between small-step AAM and big-step ADI by developing a
series of systematic transformations. Those transformations are summarized here, with their
associated section:

\begin{itemize}
\item We show how \textbf{linearization} (Section~\ref{linear}) can be used to...
\item In Section~\ref{fusing}, we then present the \textbf{fusing} transformation, and present...
\item Section~\ref{disen} discusses \textbf{disentangling}, which shows...
\item \textbf{Refunctionalization} is shown in Section~\ref{refunc}, and brings a new understanding of pushdown control-flow analysis. In this section, we present an abstract interpreter obtained from refunctionalizing a small-step abstracting abstract machine which naturally exhibits the pushdown control-flow property (as refunctionalization requires abstract execution to be sequentialized).
\item The last transformation we show is that of transforming a definitional abstract interpreter to a \textbf{direct-style} interpreter (Section~\ref{directstyle}).

These transformations are used throughout the paper, with refunctionalization and
defunctionalization of abstract interpreters playing important roles for the call stack of
the analyzed language (i.e., Scala)\note{right?}. All of this insight allows the revelation of
precisely \emph{why} small-step AAM does not have the pushdown control-flow property
(without extra effort), as well as explains why ADI naturally extends pushdown control-flow
property from its defining language.

We complete this paper by discussing related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.

% \subsection{Contributions}

@@

% We summarize our contributions as follows:

% \begin{itemize}
% \item The main contribution of this paper is the filling of the gap
%   between small-step AAM and big-step ADI by developing a series of
%   systematic transformations.
%   We show that the correspondence between abstract
%   machines and interpreters exists not only in concrete semantics artifacts,
%   but also in abstract semantics artifacts.
%   Galois connection?
%   \note{JD: Section number?}

% \item A new understanding of pushdown control-flow analysis by
%   refunctionalization.
%   We present an abstract interpreter obtained from refunctionalizing
%   a small-step abstracting abstract machine.
%   This abstract interpreter naturally exhibits the pushdown control-flow property,
%   as the refunctionalization forces us to sequentialize the
%   abstract execution.
%   We also analyze its correctness, soundness, and termination.

% \item Further, we show that defunctionalization and refunctionalization
%   of abstract interpreters play important roles for the call stack of the analyzed language.
%   Given this insight, we reveal precisely \emph{why} small-step AAM has no pushdown
%   control-flow property (without extra efforts), as well as explain
%   why ADI naturally extends pushdown control-flow property from its
%   defining language.

% \end{itemize}

% \subsection{Outline}

% \note{TR: can this be merged with contributions? (they currently lack section refs)}
% The remainder of this paper is organized as follows.
% We begin by describing the target language we will analyze in Section~\ref{background},
% including a concrete CESK machine as small-step operational semantics for it.
% We continue with a review of abstracting abstract machines (AAM) as an abstract interpretation
% counterpart of CESK machines. Following this, Section~\ref{linear} shows a conversion from the
% nondeterministic choices of closures to another layer of continuations,
% therefore linearizing the abstract execution.
% Section~\ref{fusing} then discusses fusing the driver and step functions to a single
% function.
% We further identify in Section~\ref{disen} the first-order data type representations of
% continuations, and disentangle the code blocks for dispatching data types to separate
% functions.
% Section~\ref{refunc} then represents the refunctionalized abstract abstract machines.
% In this section, the continuations represented by data types become higher-order
% functions.
% In Section~\ref{directstyle}, we transform the refunctionalized AAM to two
% different direct styles: one implemented with monads, the other using
% delimited controls.

% Finally, we discuss related work in Section~\ref{sec:related}, followed by concluding thoughts in Section~\ref{sec:conclusion}.
